### 2020-03-24 tex/gitlinks.tex

\newcommand{\codelink}[1]{\href{https://github.com/user/repo/blob/cfda45841cf4235e82cf705ab27251bd783bdf8c/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/user/repo/blob/cfda45841cf4235e82cf705ab27251bd783bdf8c/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/user/repo/blob/cfda45841cf4235e82cf705ab27251bd783bdf8c/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/user/repo/_build}{#1}}