### 2020-03-24 tex/ms.tex

% !TeX root = ./ms.tex
\documentclass[modern]{aastex62}

% Load the corTeX style definitions
\input{cortex}

% Bibliography stuff
\bibliographystyle{aasjournal}

% Begin!
\begin{document}

% Title
\title{Title here}

% Author list
\author{Fran Bartoli\'c}
\email{fbartolic@flatironinstitute.org}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}
%

\begin{abstract} 
Abstract here.
%
\href{https://github.com/rodluger/starrynight}{\color{linkcolor}\faGithub}
\end{abstract}

%
\section{Introduction}
%
Intro here \citep{Luger2019}.
Here's an equation with a proof:
%
\begin{proof}{dummy}
    \label{eq:dummy}
    1 + 1 = 2.
\end{proof}
The proof is a Jupyter notebook with a formal derivation of the solution,
an informal justification, a numerical validation, or whatever you want it
to be.
%
And here's a figure with a link:
%
\begin{figure}[h!]
    \begin{centering}
    \includegraphics[width=0.5\linewidth]{figures/dummy.pdf}
    \oscaption{pretty_function}{%
        This is a plot of a pretty function. And at the end of this
        caption is a symbol with a link to the \emph{exact} script
        that generated it, hosted on \textsf{GitHub}.
        \label{fig:dummy}
    }
    \end{centering}
\end{figure}

% Bibliography
\bibliography{bib}

\end{document}