### 2020-04-21 paper/gitlinks.tex
Added abstract and sections for the paper

\newcommand{\codelink}[1]{\href{https://github.com/fbartolic/volcano/blob/c5ab3bfe105a70f2b0b6059d49a00231b3c60efb/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/fbartolic/volcano/blob/c5ab3bfe105a70f2b0b6059d49a00231b3c60efb/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/fbartolic/volcano/blob/c5ab3bfe105a70f2b0b6059d49a00231b3c60efb/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/fbartolic/volcano/_build}{#1}}